---
title: "Plots and tables with reskit"
vignette: >
  %\VignetteIndexEntry{Plots and tables with reskit}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
---

The aims of this vignette are to:

* illustrate usage patterns of some of the functions in {reskit}
* provide a set of test/example outputs of tables and charts derived from NHP
  demand model results data.

The latter may be visually compared to / validated against visualisations
currently produced by the NHP Outputs app.

Firstly we load the functions in the {reskit} package.
Then read in a set of results in the "new" format (Parquet files); in this
vignette we just read in three of the results tables, specified using the
`tables` argument to `read_results_parquet_files()`, as these are all we will
need for the visualisations.


```{r}
#| label: setup
library(reskit)
# purrr::walk(dir(here::here("R"), full.names = TRUE), source)
```


```{r}
#| label: get-data
#| cache: true
token <- azkit::get_auth_token()
results_container <- azkit::get_container(
  Sys.getenv("AZ_RESULTS_CONTAINER"),
  token = token
)

results_path <- azkit::read_azure_table(
  Sys.getenv("AZ_RUNS_TABLE"),
  token = token,
  filter = "PartitionKey eq 'national' and app_version eq 'v5.1'",
  select = "aggregated_results_path"
) |>
  dplyr::pull("aggregated_results_path") |>
  grepl(pattern = "NDG2\\-zero") |>
  dplyr::first() # ensure we only get 1 path!

results_tables <- c(
  "default",
  "tretspef+los_group",
  "step_counts",
  "sex+age_group",
  "sex+tretspef_grouped"
)

results <- read_results_parquet_files(results_container, results_path)

default_tbl <- results[["default"]]
```

```{r}
#| label: create-dummy-data
create_dummy_default_tbl <- function(seed) {
  set.seed(seed)
  ae_pods <- paste0("aae_type-0", seq(5))
  ae_measures <- c("ambulance", "walk-in")
  ip_pods <- paste0(
    "ip_",
    c(
      paste0(c("elective", "maternity", "non-elective"), "_admission"),
      paste0("regular_", c("night", "day"), "_attender")
    )
  )
  ip_measures <- c("admissions", "beddays", "procedures")
  op_pods <- paste0("op_", c("first", "follow-up", "procedure"))
  op_measures <- paste0(c("", "tele_"), "attendances")
  ae_tbl <- tidyr::expand_grid(pod = ae_pods, measure = ae_measures)
  ip_tbl <- tidyr::expand_grid(pod = ip_pods, measure = ip_measures)
  op_tbl <- tidyr::expand_grid(pod = op_pods, measure = op_measures)
  init_tbl <- dplyr::bind_rows(ae_tbl, ip_tbl, op_tbl) |>
    dplyr::mutate(sitetret = "site1", .before = "measure")
  init_tbl2 <- dplyr::mutate(init_tbl, sitetret = "site2")
  activ_min <- 5e3L
  activ_max <- 85e3L
  site1_baselines <- sample(seq(activ_min, activ_max, 1e3L), nrow(init_tbl))
  site2_baselines <- as.integer(site1_baselines / 2)
  mitigations <- rnorm(nrow(init_tbl), 1, 0.1) *
    c(rep(0.9, nrow(ae_tbl)), rep(0.85, nrow(ip_tbl)), rep(0.8, nrow(op_tbl)))
  site1_baseline <- init_tbl |>
    dplyr::mutate(model_run = 0L, value = site1_baselines)
  site2_baseline <- init_tbl2 |>
    dplyr::mutate(model_run = 0L, value = site2_baselines)
  site1_horizons <- init_tbl |>
    dplyr::mutate(
      value = purrr::map2(site1_baselines, mitigations, generate_horizon_vals)
    )
  site2_horizons <- init_tbl2 |>
    dplyr::mutate(
      value = purrr::map2(site2_baselines, mitigations, generate_horizon_vals)
    )
  horizons_tbl <- dplyr::bind_rows(site1_horizons, site2_horizons) |>
    tidyr::unnest_longer("value", indices_to = "model_run")
  list(site1_baseline, site2_baseline, horizons_tbl) |>
    dplyr::bind_rows() |>
    dplyr::arrange(dplyr::pick(c("pod", "sitetret", "measure", "model_run")))
}

generate_horizon_vals <- function(baseline, mitigation, picks = 64) {
  as.integer(round(rnorm(picks, 1, 0.1) * baseline * mitigation))
}

create_dummy_default_tbl(seed = 1871)
```


Having shown one way of accessing the results parquet files, this vignette now
runs through the suggested pipelines of functions that generate the NHP Outputs
app tables and charts, which usually comprise in their simplest form, a data
preparation function which passes data into a table or chart creation function.

This series of example outputs is followed by a breakdown of one of the data
preparation functions, showing how a results data table may be filtered,
enhanced, summarised and arranged using {reskit} functions to produce standard
statistics and outputs.


## Outputs: tables and charts


### "Main summary table"

This uses the "default" table (from `default.parquet`), passing it to the
`compile_principal_pod_data` function which prepares the data for the table.

The only relevant filter here is for site selection.

Then the `make_principal_pod_table` function generates the table.


```{r main summary table}
results |>
  purrr::pluck("default") |>
  compile_principal_pod_data(sites = NULL) |>
  make_principal_pod_table()
```


### "Principal LoS table" - bed days measure


```{r principal los table}
results |>
  purrr::pluck("tretspef+los_group") |>
  compile_principal_los_data(measure = "Beddays", sites = NULL) |>
  make_principal_los_table()
```


### "Principal LoS table" - admissions measure

```{r}
results |>
  purrr::pluck("tretspef+los_group") |>
  compile_principal_los_data(measure = "Admissions", sites = NULL) |>
  make_principal_los_table()
```


### Impact of changes - overall ("waterfall") chart with baseline

Selecting `pods = NULL` would use data

```{r}
results |>
  purrr::pluck("step_counts") |>
  compile_change_factor_data(
    measure = "admissions",
    activity_type = "ip",
    pods = c(
      "Daycase Admission",
      "Maternity Admission",
      "Non-Elective Admission"
    ),
    sites = NULL,
    include_baseline = TRUE
  ) |>
  make_overall_cf_plot()
```



### Impact of changes - overall ("waterfall") chart without baseline

```{r}
results |>
  purrr::pluck("step_counts") |>
  compile_change_factor_data(
    activity_type = "Inpatient",
    pods = c(
      "Daycase Admission",
      "Maternity Admission",
      "Non-Elective Admission"
    ),
    measure = "Admissions",
    sites = NULL,
    include_baseline = FALSE
  ) |>
  make_overall_cf_plot()
```


### Impact of changes - individual TPMA chart, Admissions measure, sort by value


```{r}
measure <- "Admissions"

results |>
  purrr::pluck("step_counts") |>
  compile_indiv_change_factor_data(
    activity_type = "Inpatient",
    pods = c(
      "Daycase Admission",
      "Maternity Admission",
      "Non-Elective Admission"
    ),
    measure = measure,
    sort_by = "value",
    sites = NULL
  ) |>
  make_individual_cf_plot(x_axis_label = measure)
```


### Impact of changes - individual TPMA chart, Admissions measure, sorted by name


```{r}
measure <- "Admissions"

results |>
  purrr::pluck("step_counts") |>
  compile_indiv_change_factor_data(
    activity_type = "Inpatient",
    pods = c(
      "Daycase Admission",
      "Maternity Admission",
      "Non-Elective Admission"
    ),
    measure = measure,
    sort_by = "tpma_label",
    sites = NULL
  ) |>
  make_individual_cf_plot(x_axis_label = measure)
```


### Impact of changes - individual TPMA chart, Bed Days measure, sort by value


```{r}
measure <- "Beddays"

results |>
  purrr::pluck("step_counts") |>
  compile_indiv_change_factor_data(
    activity_type = "Inpatient",
    pods = c(
      "Daycase Admission",
      "Maternity Admission",
      "Non-Elective Admission"
    ),
    measure = measure,
    sort_by = "value",
    sites = NULL
  ) |>
  make_individual_cf_plot(x_axis_label = measure)
```


### Activity in detail table - IP activity by sex and age group


```{r}
results |>
  compile_detailed_activity_data()
# filter_detailed_activity_data(
#   aggregation = "age_group",
#   activity_type = "ip",
#   pods =
# )
```






## Breakdown of a data preparation function

This section shows the interim steps generated by the constituent functions and
processes with an example data preparation function.

We will look at `compile_principal_pod_data()` here.

```{r}
#| echo: false
def <- results[["default"]]
```

Starting with the default table, before applying any functions to it, we can see
that it has `r ncol(def)` columns.
Several of these are "metadata" columns - containing only a single unique value,
such as the name of the scenario - that we won't need for the further processing
of the table.

The data columns we need to work with are:

* `sitetret`
* `pod`
* `measure`
* `model_run`, and
* `value`

(Other results tables will have a different selection of column names that we
would need to use.)


```{r}
results[["default"]]
```

In preparing the data for visualisation, we apply the
`compile_principal_pod_data()` function.
This runs through a sequence of functions that transform and summarise the data.
These are, in turn:

* `prepare_sites_principal_pod_data()`
* `filter_to_selected_sites()` + `summarise_for_all_sites()`
* `add_change_cols()`
* final preparations, via for example `forcats::fct()` and `dplyr::arrange()`


```{r}
interim1 <- results[["default"]] |>
  prepare_sites_principal_pod_data()
interim1
```

The `prepare_sites_principal_pod_data()` function applies some filters to the
data table and adds a human-friendly `pod_label` column (via `relabel_pods()`),
then - even more importantly - applies `calculate_principal_stats()`.
This last function calculates the "p10" and "p90" values as well as the mean of
the values across all the model runs for each combination of `pod` and
`measure` (in this case).

### Breaking down `prepare_sites_principal_pod_data()`

Let's have a look at how these functions change the shape and content of the
data table to create the `interim1` object shown above.

Firstly there are some initial changes and filters:

```{r}
interim1a <- results[["default"]] |>
  dplyr::filter(dplyr::if_any("measure", \(x) x %in% keep_measures())) |>
  dplyr::filter(
    # exclude outpatient procedures from tele-attendances count only
    dplyr::if_any("measure", \(x) x != "tele_attendances") |
      dplyr::if_any("pod", \(x) x != "op_procedure")
  ) |>
  dplyr::mutate(dplyr::across("pod", \(x) sub("^aae.*$", "aae", x)))
interim1a
```

Now joining to a lookup of POD (point of delivery) codes and labels:

```{r}
interim1b <- interim1a |>
  inner_join_for_labels(get_principal_pods()) |>
  relabel_pods()
interim1b
```

We now have a new `activity_type_label` column.

And then, summarising by certain fields to calculate the mean values:


```{r}
interim1c <- interim1b |>
  calculate_principal_stats(default_group_cols("activity_type_label")) |>
  dplyr::filter(dplyr::if_any("stat", \(x) x == "mean")) |>
  dplyr::select(!"stat")
interim1c
```

```{r}
identical(interim1c, interim1)
```

`calculate_principal_stats()` helps us treat model run 0 as a "baseline" value,
and model runs 1-256 as the basis for our "principal" projection.

We now have a table with just `r nrow(interim1c)` rows and `r ncol(interim1c)`
columns, and some quite different column names to the table we started with.

### Subsequent steps

The above steps will be the same no matter what the user of the NHP Outputs app
selects.

Next there is a further step that will depend on the selection of sites chosen
by the user.
The default selection is `NULL`, meaning all sites in the data will be included.
The dataset we are using in this vignette is "national", which means we don't
have any specific sites to filter by in any case.
But for a local scheme, a user could choose to filter results data just to
specific sites.

The `filter_to_selected_sites()` and `summarise_for_all_sites()` functions
deal with this.

`add_change_cols()` calculates the change from the baseline values to the
mean principal projected values, as an absolute number and as a percentage.

Let's apply these anyway to our output from
`prepare_sites_principal_pod_data()`, and then apply `add_change_cols()`.


```{r}
interim2 <- interim1 |>
  filter_to_selected_sites(NULL) |>
  summarise_for_all_sites(site_col = "sitetret") |>
  add_change_cols()
interim2
```

Then our final steps in this breakdown of `compile_principal_pod_data()`.

We should have ended up reproducing the output we get from simply running
`compile_principal_pod_data(results[["default"]])`:

```{r}
at_levels <- c(
  "Inpatient Admissions",
  "Inpatient Bed Days",
  "Outpatient",
  "A&E"
)
interim3 <- interim2 |>
  dplyr::mutate(
    dplyr::across("activity_type_label", \(x) forcats::fct(x, at_levels)),
    # display pods in descending order of baseline value, by activity type
    dplyr::across("pod_label", \(x) {
      forcats::fct_reorder(x, .data[["baseline"]], sum, .desc = TRUE)
    })
  ) |>
  dplyr::arrange(dplyr::pick(c("activity_type_label", "pod_label")))
interim3
```

This data table is now in the expected format to be passed through to
`make_principal_pod_table()`.

### Other data preparation functions

Exploring each preparation pipeline in detail would make this a very lengthy
vignette.
However they do share some common functions and techniques.

Other data preparation functions in {reskit} include:

* `compile_principal_los_data()`, used to prepare data for `make_principal_los_table()`
* `compile_change_factor_data()`, used to prepare data for `make_overall_cf_plot()`
* and so on
