---
title: "Plots and tables with reskit"
vignette: >
  %\VignetteIndexEntry{Plots and tables with reskit}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
lang: en-GB
date: today
theme: lumen
code-fold: show
fig-width: 12
fig-asp: 0.618
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
---

The aims of this vignette are to:

* illustrate usage patterns of some of the functions in {reskit}
* provide a set of test/example outputs of tables and charts derived from NHP
  demand model results data.

The latter may be visually compared to / validated against visualisations
currently produced by the NHP Outputs app.

Firstly we load the functions in the {reskit} package.
Then we create a fake `results` object which consists of synthetic data
created by some internal reskit functions.
It ought to go without saying that the numbers in these synthetic tables are
entirely fictitious and unrealistic.
They exist purely to provide a test basis for the data processing and
visualisation functions in this package.


```{r}
#| label: setup
# library(reskit)
purrr::walk(dir(here::here("R"), full.names = TRUE), source)
seed <- 1871
```

## Example using azkit and reskit to download results

Here is an example of some code that you might use to access results data from
Azure storage


```r
token <- azkit::get_auth_token()
results_container <- azkit::get_container(
  Sys.getenv("AZ_RESULTS_CONTAINER"),
  token = token
)

# use various filters to get the right dataset (PartitionKey), model version and
# scenario from a runs lookup table
results_path <- azkit::read_azure_table(
  Sys.getenv("AZ_RUNS_TABLE"),
  token = token,
  filter = "PartitionKey eq 'national' and app_version eq 'v5.1'",
  select = "aggregated_results_path"
) |>
  dplyr::pull("aggregated_results_path") |>
  grepv(pattern = "NDG2\\-zero") |>
  dplyr::first() # ensure we only get 1 path!

results <- read_results_parquet_files(results_container, results_path)
```

```{r}
#| label: create-demo-data
#| cache: true
#| cache-refresh: false
default_tbl <- create_demo_default_tbl(seed)
tretspef_losgroup_tbl <- create_demo_tretspef_losgroup_tbl(seed)
sex_agegroup_tbl <- create_demo_sex_agegroup_tbl(seed)
sex_tretspef_tbl <- create_demo_sex_tretspef_tbl(seed)
stepcounts_tbl <- create_demo_stepcounts_tbl(seed)

# fmt: skip
tbl_names <- c(
  "default", "tretspef+los_group", "sex+age_group",
  "sex+tretspef_grouped", "step_counts"
)

results <- list(
  default_tbl,
  tretspef_losgroup_tbl,
  sex_agegroup_tbl,
  sex_tretspef_tbl,
  stepcounts_tbl
) |>
  rlang::set_names(tbl_names)
```


Having shown one way of accessing the results parquet files, this vignette now
runs through the suggested pipelines of functions that generate the NHP Outputs
app tables and charts, which usually comprise in their simplest form, a data
preparation function which passes data into a table or chart creation function.

This series of example outputs is followed by a breakdown of one of the data
preparation functions, showing how a results data table may be filtered,
enhanced, summarised and arranged using {reskit} functions to produce standard
statistics and outputs.


## Outputs: tables


### "Main summary table" - overall change by Point of Delivery

This uses the "default" table (from `default.parquet`), passing it to the
`compile_principal_pod_data` function which prepares the data for the table.

The only relevant filter here is for site selection.

Then the `make_principal_pod_table` function generates the table.


```{r}
#| label: main-summary-table1
results |>
  compile_principal_pod_data() |>
  make_principal_pod_table()
```

Now the same thing again but testing the site selection feature.
We should get a similar style table but with smaller numbers as we have
filtered down to just one site.

```{r}
#| label: main-summary-table2
results |>
  compile_principal_pod_data(sites = "site1") |>
  make_principal_pod_table()
```


### The principal LoS table


This uses data from the "tretspef+los_group" results table.

With `measure` set to "beddays":

```{r}
#| label: los-summary-table1
results |>
  compile_principal_los_data(measure = "beddays") |>
  make_principal_los_table()
```


With `measure` set to "admissions":

```{r}
#| label: los-summary-table2
results |>
  compile_principal_los_data(measure = "admissions", sites = "site2") |>
  make_principal_los_table()
```

## Outputs: charts

### Impact of changes charts

For these we need to use carefully chosen combinations of measure, activity
type, and pod.

#### Overall ("waterfall") chart with baseline

As this is for admissions, "efficiencies" should not be shown on the y-axis.

```{r}
#| label: waterfall-chart1
results |>
  compile_change_factor_data(
    measure = "admissions",
    activity_type = "ip",
    pods = c(
      "ip_elective_admission",
      "ip_maternity_admission",
      "ip_non-elective_admission"
    ),
    include_baseline = TRUE
  ) |>
  make_overall_cf_plot()
```


```{r}
#| label: waterfall-chart2
results |>
  compile_change_factor_data(
    measure = "beddays",
    activity_type = "ip",
    pods = c(
      "ip_elective_admission",
      "ip_maternity_admission",
      "ip_non-elective_admission"
    ),
    sites = "site1",
    include_baseline = TRUE
  ) |>
  make_overall_cf_plot()
```

Let's look at an outpatients chart...

```{r}
#| label: waterfall-chart3
results |>
  compile_change_factor_data(
    measure = "attendances",
    activity_type = "op",
    pods = c("op_first", "op_follow-up")
  ) |>
  make_overall_cf_plot()
```


...and one without the baseline:

```{r}
#| label: waterfall-chart4
results |>
  compile_change_factor_data(
    measure = "attendances",
    activity_type = "op",
    pods = c("op_first", "op_follow-up"),
    include_baseline = FALSE
  ) |>
  make_overall_cf_plot()
```


And a couple of individual change factors charts:



```{r}
#| label: indiv-waterfall-chart1
results |>
  compile_indiv_change_factor_data(
    measure = "attendances",
    activity_type = "op",
    pods = NULL
  ) |>
  make_individual_cf_plot()
```
